---
title: "LaBella_03"
author: "Rachel LaBella"
date: "September 22, 2016"
output: html_document
---
```{r, echo = TRUE}
library(readr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(forcats)
```
####1) Complete problems 10 and 17-18 on pg. 109-111. Use R where possible. Data sets (so you don’t have to type things in) are available at http://www.zoology.ubc.ca/~whitlock/ABD/teaching/datasets.html.

####2.1) Load the data using readr and make the Month_Names column into a factor whose levels are in order of month.
``` {r, echo = TRUE}
sea_ice <- read_csv("../sea_ice.csv")
sea_ice <- sea_ice %>%
  mutate(month_name = fct_inorder(month_name))
levels(sea_ice$month_name)
```

####2.2) Make a boxplot showing the variability in sea ice extent every month.
``` {r, echo = TRUE}
sea_ice_plot <- ggplot(sea_ice, mapping = aes(x = month_name, y = extent)) 
sea_ice_plot +
  geom_boxplot()
```


####2.3) Use dplyr to get the annual minimum sea ice. Plot minimum ice by year, and add a trendline (either a smooth spline or a straight line).

####2.4) Plot sea ice by year, with different lines for different months. Then, add a new column (mutate!) using the ggplot2 function cut_interval(Month, n=4) - this will create four even bins. Seasons, if you will. Use facet_wrap on the same plot and split the plot into seasons.

####2.5) Last, make a line plot of sea ice by month. Gussy it up with colors by year, a different theme, and whatever other annotations, changes to axes, etc., you think best show the story of this data.

####2.6 Extra Credit) Make it animated.